### 2025-08-23 evaluating-prediction-accuracy.qmd
20250822 - convert caret to tidymodels
To address https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues/10

```{r}
library("petersenlab")
library("pROC")
library("magrittr")
library("viridis")
library("viridisLite")
library("msir")
library("caret")
library("tidyverse")
```

@@

For example, here is an example of estimating predictive $R^2$ using 10-fold cross-validation using the `caret` package [@R-caret].

@@

# Set up 10-fold cross-validation
train_control <- caret::trainControl(
  method = "cv",
  number = 10)

# Fit the model with cross-validation
model_cv <- caret::train(
  outcomeVariable ~ predictorVariable1 + predictorVariable2,
  data = mydata,
  method = "lm",
  na.action = na.exclude,
  trControl = train_control,
  metric = "Rsquared")

# View R-squared across the folds
model_cv$results$Rsquared
```

Here is an example of estimating predictive $R^2$ with a training and testing data set using the `caret` package [@R-caret].

inTrain <- caret::createDataPartition(
  y = as.factor(mydata$outcomeVariable),
  p = .75, # the proportion of data for the training data set
  list = FALSE
)

training <- mydata[ inTrain,]
testing  <- mydata[-inTrain,]

training_model <- lm(
  outcomeVariable ~ predictorVariable1 + predictorVariable2,
  data = training)

testing_predictions <- predict(
  training_model,
  newdata = testing)

petersenlab::accuracyOverall(
  predicted = testing_predictions,
  actual = testing$outcomeVariable)$rsquared
```

>>>>

```{r}
library("petersenlab")
library("pROC")
library("magrittr")
library("viridis")
library("viridisLite")
library("msir")
library("tidymodels")
library("tidyverse")
```

@@

For example, here is an example of estimating predictive $R^2$ using [10-fold cross-validation](#sec-kFoldCrossValidation) using the `tidymodels` ecosystem of packages [@R-tidymodels; @Kuhn2020_packages].

```{r}
# Set seed for reproducibility
set.seed(52242)

# Set up Cross-Validation
folds_kFold <- rsample::group_vfold_cv(
  projectionsWithActuals_seasonal,
  group = player_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
  v = 10) # 10-fold cross-validation

folds <- folds_kFold

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints ~ raw_points, # use all predictors
  data = projectionsWithActuals_seasonal)

# Define Model
lm_spec <- parsnip::linear_reg() %>%
  parsnip::set_engine("lm") %>%
  parsnip::set_mode("regression")

# Workflow
lm_wf <- workflows::workflow() %>%
  workflows::add_recipe(rec) %>%
  workflows::add_model(lm_spec)

# Fit Model with Cross-Validation
cv_results <- tune::fit_resamples(
  lm_wf,
  resamples = folds,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_resamples(save_pred = TRUE)
)

# View Cross-Validation metrics
tune::collect_metrics(cv_results)

# Fold-level R-squared
tune::collect_metrics(cv_results, summarize = FALSE) %>%
  filter(.metric == "rsq") %>%
  select(id, .estimate)
```

Here is an example of estimating predictive $R^2$ with a training and hold-out (testing) data set:

```{r}
# Set seed for reproducibility
set.seed(52242)

# Create 75/25 split
split <- rsample::group_initial_split(
  projectionsWithActuals_seasonal,
  group = player_id,
  prop = 0.75)

# Fit model on training, evaluate on testing
last_fit_results <- tune::last_fit(
  lm_wf,
  split = split,
  metrics = yardstick::metric_set(rmse, mae, rsq)
)

# View predictive R-squared and other metrics
tune::collect_metrics(last_fit_results)
```

Or, alternatively:

@@

# Set seed for reproducibility
set.seed(52242)

# Create 75/25 split
split <- rsample::group_initial_split(
  projectionsWithActuals_seasonal,
  group = player_id,
  prop = 0.75)

training <- rsample::training(split)
testing  <- rsample::testing(split)

# Fit model on training set
training_model <- parsnip::fit(
  lm_wf,
  data = training)

# Predict on test set
predictions <- predict(
  training_model,
  new_data = testing) %>%
  bind_cols(testing)

# Predictive R-squared
yardstick::rsq(
  predictions,
  truth = fantasyPoints,
  estimate = .pred)
```
